Automated Formal Synthesis of Provably Safe Digital Controllers for Continuous Plants

### 2019-07-08 automatica/response.tex

\vspace{2em}
{\bf Comment: } {\itshape he results are evaluated on about a dozen and a half examples from the controls literature all with relatively small
dimensionality (2 to 5 real variables in the continuous state space), and while the authors claim the software artifacts and benchmarks are available
for evaluation, when this reviewer browsed to the link (http://www.cprover.org/DSSynth/), a page with "TBD" was shown. I did a little digging and
eventually found what I presume is the same / similar artifacts here: https://github.com/ssvlab/dsverifier and/or http://dsverifier.org/}

\vspace{1em}
{\bf Response: }

Fixed this now

\todo[inline]{Elizabeth: fix this}
\vspace{2em}
{\bf Comment: } {\itshape On the control theory side, the approach uses standard results to synthesize stabilizing controllers for LTI systems,
specifically linear state feedback controllers (u=Kx with a possible reference signal to track), under the assumption the plant is controllable. It
seems the approach requires full-state observability as well, which is a fairly standard assumption as one can typically implement various filters /
estimators (Kalman, Leuenberger, etc.) to ensure state observability assuming detectability of the plant, but the authors should clarify this point to
make the assumptions more precise.}

\vspace{1em}
{\bf Response: }
We do need full observability of the (states of the) model, and have clarified this on page [X] as follows: \ldots

@@

\vspace{1em}
{\bf Response: }

\vspace{2em}
{\bf Comment: }
{\itshape Finally, while checking the $2^n$ vertices
are sufficient (theorem 1), I doubt this is necessary, as e.g. the star set approaches avoid this blow-up that would also limit the dimensionality
the approach can work toward.}
\vspace{1em}

{\bf Response: }
\todo[inline]{Alessandro: response to this (Theorem 1)}

\rev{It is probably true that the check on all vertices might not be necessary and could be simplified in special cases or under some conditions. The check however, as it currently is implemented, does not represent the bottleneck of our overall method, so we can safely reply on it as it stands. }

@@

It is easy to see that after two transformations, the initial set would be
\begin{verbatim}
-0.25 <= x <= 0 and -0.25 <= y <= 0.
\end{verbatim}
The unsafe set lies completely within the reachable set, but none of the edges in the hyperrectangle is unsafe. The next time the initial set reaches
this quadrant, the reachable set is
\begin{verbatim}
-1/(2^6) <= x <= 0
-1/(2^6) <= y <= 0.
\end{verbatim}

I believe that for this case, the K returned from completeness is 4 or 5.

The same theorem is also present in [1]. The proof, upon careful inspection, is not a proof of the statement given in the theorem. It only proves
that the reachable set is the convex set obtained by the trajectories of the corner states. However, it does not prove that if the trajectories are
safe, their convex combination is also safe. In fact, as given in the above counterexample, the vertices of polytope are safe does not mean that the
convex combination is safe.

I think this bug needs to be fixed in the BMC part of the verifier. Instead of checking the individual trajectories, you should in fact compute the
convex combination of trajectories (reachable set) and then perform the verification.}

\vspace{1em}
{\bf Response: }
\todo[inline]{Alessandro: Response to this (Theorem 1)}
\vspace{2em}

\rev{We have emphasised that the initial set is assumed to lie within the safe set - the contrary would trivially result in an unsafe condition. 
Recall further that the sets are hyper-boxes, thus convex. In view of the linear dynamics of the considered models, in practice both are assumed to contain the origin.  }

### 2019-07-09 automatica/response.tex

\setlength\parindent{0pt}

\usepackage{amssymb,latexsym,amsfonts,amsmath, amsthm,color,mathrsfs}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{todonotes}
%\usepackage{utf8}
\usepackage{fullpage}
\def\SS#1{{\textcolor{red}{ {\bf SS:} #1}}}
\renewcommand{\b}{\textcolor{blue}}
%\newcommand{\red}{\textcolor{red}}
%
\newcommand{\vold}[1]{\colorbox{red!20}{\bfseries #1}}
\newcommand{\vnew}[1]{\colorbox{blue!30}{\bfseries #1}}

\renewcommand{\t}{\new{(t)}}
\newcommand{\tp}{\new{(t+1)}}
\newcommand{\fei}[1]{{\color{purple}#1 \color{black}}}
\newcommand{\rev}[1]{\textcolor{blue}{#1}}

\begin{document}
\begin{center}
{\Large Responses to the Reviews}\\\vspace{0.3cm}
\today
\end{center}
\vspace{1cm}
Dear Associate Editor and Reviewers,\vspace{0.3cm}
\\
We would like to deeply thank you for reviewing our paper and providing us with the constructive comments and suggestions. We have carefully addressed
all the comments and incorporated them in the revised manuscript.

We present below a detailed discussion of all comments, questions, and concerns raised by the reviewers. The major modifications are highlighted in
blue in the revised manuscript. For clarity,  we also use \vold{v1} and \vnew{v2} in the response letter for referring respectively to the previous
and current versions of the manuscript.
%I do not like color coding that much. And definitely not with red and blue.
\\
\\
With kind regards,\vspace{0.1cm}
\\
The Authors
%
\newpage

@@

{\bf Comment: }{\itshape This paper presents an approach for synthesizing 
digital/software controllers for continuous plants (linear time invariant / LTI) 
such that they satisfy safety/stability requirements, as would show up in closed-loop 
control of cyber-physical systems (CPS). The approach extends earlier work of the authors, 
the primary of which in my view is toward floating-point implementation as opposed to only fixed-point. 
I did not entirely buy that using CEGIS is a contribution, as from looking over the earlier related publications 
(e.g., ASE'17, CAV'17, HSCC'17, SYNT'18 abstract), all of them use CEGIS.}

\vspace{1em}

{\bf Response: }

This journal submission encompasses and summarises the mentioned conference articles (CAV, HSCC, ASE) or workshop abstract (SYNTH). 
It is a more mature, complete, and comprehensive presentation of the conference material. 
Of these conference articles, the CAV/HSCC contributions presented the theory underlying the approach, 
whereas the ASE is a tool paper. 
The application of the CEGIS architecture to the synthesis of digital controllers for continuous plants is indeed a novel contribution of this line of work.

\todo[inline]{Do we actually say that CEGIS is a contribution? We should change that. Check that we don't say CEGIS is a contribution. AA: I think the reviewers complain about our earlier conference publications. We should rectify this view. I've added the text above. }

### 2019-07-09 automatica/response.tex

\vspace{2em}
{\bf Comment: } {\itshape he results are evaluated on about a dozen and a half examples from the controls literature all with relatively small
dimensionality (2 to 5 real variables in the continuous state space), and while the authors claim the software artifacts and benchmarks are available
for evaluation, when this reviewer browsed to the link (http://www.cprover.org/DSSynth/), a page with "TBD" was shown. I did a little digging and
eventually found what I presume is the same / similar artifacts here: https://github.com/ssvlab/dsverifier and/or http://dsverifier.org/}

\vspace{1em}
{\bf Response: }

Fixed this now

\todo[inline]{Elizabeth: say what the problem was, fix this if possible, then thank the reviewer for pointing this out. }

\vspace{2em}
{\bf Comment: } {\itshape On the control theory side, the approach uses standard results to synthesize stabilizing controllers for LTI systems,
specifically linear state feedback controllers (u=Kx with a possible reference signal to track), under the assumption the plant is controllable. It
seems the approach requires full-state observability as well, which is a fairly standard assumption as one can typically implement various filters /
estimators (Kalman, Leuenberger, etc.) to ensure state observability assuming detectability of the plant, but the authors should clarify this point to
make the assumptions more precise.}

\vspace{1em}
{\bf Response: }
We do need full observability of the (states of the) model, and have clarified this in Section 3.1, as follows: {\it We assume full observability of the states of the model, namely the output states correspond with the model variables.  } We recall this assumption later in the text.

\vspace{2em}
{\bf Comment: } {\itshape Theorem 2 states completeness of the approach, specifically that there exists a finite k for which BMC can be done up-to to
provide a guarantee of future safety for all larger k' greater than k. Assuming the benchmarks all work up to some completeness threshold, I would like to see
the completeness threshold for each benchmark in table 1. The reason being, I find this result a bit surprising in the given context (albeit not
unbelievable given the focus on LTI) and would like to see further evidence illustrating the point beyond the fairly informal proof, and it would
greatly enhance the existing discussion of timeouts, etc. related to completeness thresholds being too high for some cases.}

### 2019-07-09 automatica/paper.tex
response

We consider models of physical plants expressed as ordinary differential
equations, which we assume are controllable~\cite{Astrom08}: 
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), 
\end{align}
%
where 
$\vec{x} \in \mathbb{R}^n$,  
$\vec{u} \in \mathbb{R}^p$, 
$\mat{A} \in \mathbb{R}^{n \times n}$, 
$\mat{B} \in \mathbb{R}^{n \times p}$,
and $t \in \mathbb R_0^+$ denotes continuous time. 
%$\mat{A}$ and $\mat{B}$ are matrices that fully characterize the dynamics of the continuous plant,
We denote with $x(0)$ the model initial condition, which can be non-deterministic.

@@

We consider state feedback control architectures, 
where $u_k$ (notice we work with the discretized signal) is $u_k = r_{k} - K x_k$. 
Here $K \in \mathbb{R}^{p \times n}$ is a state-feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital). 
We will assume $r_k=0$, thus obtaining the closed-loop model 
$\vec{x}_{k+1}=(\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_k$.

@@

\begin{proof}
%
An asymptotically stable model is known to have converging dynamics.  Assume the
closed-loop matrix eigenvalues are not repeated (which is sensible to do,
since we select them).  The distance of the trajectory from the reference
point (which, for linear models, is the origin) decreases over time within subspaces related to real-valued
eigenvalues; however, this is not the case in general when dealing with
complex eigenvalues.  Consider the closed-loop matrix that updates the
states in every discrete time step, and select the eigenvalue $\vartheta$
with the smallest (non-trivial) imaginary value.  Between every pair of
consecutive time steps $k\,T_s$ and $(k+1)\,T_s$, the dynamics projected on
the corresponding eigenspace rotate $\vartheta T_s$ radians.  Thus, taking
$\overline{k}$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, after
$k{\geq}\overline{k}$ steps we have completed a full rotation, which results
in a point closer to the origin.  The synthesized $\overline{k}$ is the completeness threshold.
\qed 
%
\end{proof}

>>>>

We consider models of physical plants expressed as ordinary differential
equations, which we assume are controllable~\cite{Astrom08}: 
%
\begin{align}
\label{eq:ode1}
\dot{x}(t) = \mat{A}\vec{x}(t)+ \mat{B} \vec{u}(t), 
\end{align}
%
where 
$\vec{x} \in \mathbb{R}^n$,  
$\vec{u} \in \mathbb{R}^p$, 
$\mat{A} \in \mathbb{R}^{n \times n}$, 
$\mat{B} \in \mathbb{R}^{n \times p}$,
and $t \in \mathbb R_0^+$ denotes continuous time. 
%$\mat{A}$ and $\mat{B}$ are matrices that fully characterize the dynamics of the continuous plant,
We denote with $x(0)$ the model initial condition, which can be non-deterministic.  
\rev{ We assume full observability of the states of the model, namely the output states correspond with the model variables.  }

@@

\rev{ Recall that the states of the model are fully accessible (namely, observable). } 
We consider state feedback control architectures, 
where $u_k$ (notice we work with the discretized signal) is $u_k = r_{k} - K x_k$. 
Here $K \in \mathbb{R}^{p \times n}$ is a state-feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital). 
We will assume $r_k=0$, thus obtaining the closed-loop model 
$\vec{x}_{k+1}=(\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_k$.

@@

\begin{proof}
%
An asymptotically stable model is known to have converging dynamics.  Assume the
closed-loop matrix eigenvalues are not repeated (which is sensible to do,
since we select them).  The distance of the trajectory from the reference
point (which, for linear models, is the origin) decreases over time within subspaces related to real-valued
eigenvalues; however, this is not the case in general when dealing with
complex eigenvalues.  Consider the closed-loop matrix that updates the
states in every discrete time step, and select the eigenvalue $\vartheta$
with the smallest (non-trivial) imaginary value.  Between every pair of
consecutive time steps $k\,T_s$ and $(k+1)\,T_s$, the dynamics projected on
the corresponding eigenspace rotate $\vartheta T_s$ radians.  Thus, taking
$\overline{k}$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, after
$k{\geq}\overline{k}$ steps we have completed a full rotation, which results
in a point closer to the origin.  The synthesized $\overline{k}$ is the completeness threshold. 
\rev{[I will try to spin this more formally. ]}
\qed 
%
\end{proof}
